Search Strategies for Functional Logic Programming

### 2012-01-18 papers/atps2012/paper.tex
ATPS: reference added
Added Benchmark results for first Solution

The operational semantics of Curry is based on an optimal evaluation strategy
\cite{AntoyEchahedHanus00JACM} which is a conservative extension
of lazy functional programming and (concurrent) logic programming.
A big-step and a small-step operational semantics of Curry
can be found in \cite{AlbertHanusHuchOliverVidal05}.
Curry's semantics is sound in the sense of logic programming,
i.e., each computed result is correct and for each
correct result there is a more general computed one
\cite{AntoyEchahedHanus00JACM}.
In order to achieve completeness,
one has to take \emph{all} possible non-deterministic derivation paths
into account. In contrast to Prolog, which fixes
a (potentially incomplete) depth-first search strategy to find solutions,
Curry does not fix a particular search strategy.
Thus, different Curry implementations
can support various search strategies.
For instance, the Curry implementation PAKCS \cite{Hanus10PAKCS},
which compiles Curry programs into Prolog programs,
supports only a depth-first search strategy.
MCC \cite{Lux99FLOPS} compiles Curry programs into C programs
and uses also a depth-first search strategy to find the
solutions of a given top-level goal.
In addition, MCC offers the encapsulation of search
(see below) so that other search strategies,
like a complete breadth-first search strategy, can be used
inside a Curry program.
The Curry implementation KiCS \cite{BrasselHuch07,BrasselHuch09},
which compiles Curry programs into Haskell programs,
offers depth-first and breadth-first search strategies
for top-level goals as well the encapsulation of search
with user-definable strategies.
In this paper we consider the Curry implementation
KiCS2 \cite{BrasselHanusPeemoellerReck11},
which is based on similar ideas than KiCS but uses
a different compilation model avoiding side effects
to enable better optimizations for target programs.
KiCS2 also offers different search strategies for top-level goals
and the encapsulation of search with user-definable strategies,
which is described next.

@@

\begin{figure}[ht]
\centering
\begin{tabular}{|@{~~}l@{~~}|*{7}{@{~~}r@{~~}|}}
\hline
Program  &   DFS  & IDS(+1)& IDS(*2)&  eDFS  &  eBFS  &eIDS(+1)&eIDS(*2)\\
\hline
PermSort &        &        &        &        &        &        &        \\
Last     &        &        &        &        &        &        &        \\
Half     &        &        &        &        &        &        &        \\
\hline
\end{tabular}
\caption{Benchmarks: computing a single value}
\label{fig:bench-first-solution}
\end{figure}

>>>>

The operational semantics of Curry is based on an optimal evaluation strategy
\cite{AntoyEchahedHanus00JACM} which is a conservative extension
of lazy functional programming and (concurrent) logic programming.
A big-step and a small-step operational semantics of Curry
can be found in \cite{AlbertHanusHuchOliverVidal05}.
Curry's semantics is sound in the sense of logic programming,
i.e., each computed result is correct and for each
correct result there is a more general computed one
\cite{AntoyEchahedHanus00JACM}.
In order to achieve completeness,
one has to take \emph{all} possible non-deterministic derivation paths
into account. In contrast to Prolog, which fixes
a (potentially incomplete) depth-first search strategy to find solutions,
Curry does not fix a particular search strategy.
Actually, descriptions of the model-theoretic \cite{GonzalezEtAl99}
or operational \cite{AlbertHanusHuchOliverVidal05}
semantics of functional logic languages do not take
a search strategy into account.
Thus, different Curry implementations
can support various search strategies.
For instance, the Curry implementation PAKCS \cite{Hanus10PAKCS},
which compiles Curry programs into Prolog programs,
supports only a depth-first search strategy.
MCC \cite{Lux99FLOPS} compiles Curry programs into C programs
and uses also a depth-first search strategy to find the
solutions of a given top-level goal.
In addition, MCC offers the encapsulation of search
(see below) so that other search strategies,
like a complete breadth-first search strategy, can be used
inside a Curry program.
The Curry implementation KiCS \cite{BrasselHuch07,BrasselHuch09},
which compiles Curry programs into Haskell programs,
offers depth-first and breadth-first search strategies
for top-level goals as well the encapsulation of search
with user-definable strategies.
In this paper we consider the Curry implementation
KiCS2 \cite{BrasselHanusPeemoellerReck11},
which is based on similar ideas than KiCS but uses
a different compilation model avoiding side effects
to enable better optimizations for target programs.
KiCS2 also offers different search strategies for top-level goals
and the encapsulation of search with user-definable strategies,
which is described next.

@@

\begin{figure}[ht]
\centering
\begin{tabular}{|@{~~}l@{~~}|*{7}{@{~~}r@{~~}|}}
\hline
Program  &   DFS  & IDS(+1)& IDS(*2)&  eDFS  &  eBFS  &eIDS(+1)&eIDS(*2)\\
\hline
PermSort &  12.57 & 804.82 &  32.09 &  34.75 &  67.02 &  70.47 &  46.66 \\
Last     &   0.10 & 661.92 &   0.24 &   0.35 &   0.19 &   6.57 &   0.39 \\
Half     &   0.34 &  72.64 &   0.58 &   0.69 &   0.59 &   0.78 &   0.78 \\
Graph    &   0.00 &   0.10 &   0.04 &   0.00 &   0.04 &   0.04 &   0.03 \\
HorseMan &   2.06 & 636.62 &   6.64 &   2.52 &   2.60 &   2.61 &   2.57 \\
MAC      &   0.14 &  38.75 &   0.94 &   0.94 &   2.81 &   2.75 &   2.79 \\
Queens   &   1.40 & 370.26 &  33.65 &   2.44 &   8.95 &   8.86 &   3.04 \\
NDNums   &   oom  &  47.89 &   0.14 &    oom &   oom  &    oom &   0.47 \\
\hline
\end{tabular}
\caption{Benchmarks: computing a single value}
\label{fig:bench-first-solution}
\end{figure}
